---
title: "Tooth Growth Data Analysis"
author: "Aaron Arnold"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
  )
```

```{r}
# Questions I will be answering in this program:
# 1. Does vitamin C help in tooth growth in guinea pigs?
# 2. Which method of vitamin C intake is better, ascorbic acid or orange juice?

# Libraries needed
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)


tg <- (ToothGrowth)

ggplot(tg, aes(x=as.factor(dose), y=len, fill=supp)) + geom_boxplot() + labs(title="Tooth Growth by Dose and Supplement Type", x="Dose (mg/day)", y="Tooth Length (microns)", fill="Supplement")
```

This graph is a box plot of the original data from the ToothGrowth data set.

<br>
<br>
```{r}
# This r-block is going to be my filter.
# I will be filtering out the outliers in the ToothGrowth box plot.

Dose0.5Filter <- filter(ToothGrowth, dose == 0.5)

Dose1_VC <- filter(tg, dose == 1 & supp == "VC" & len <= 19)
Dose1_OJ <- filter(tg, dose == 1 & supp =="OJ")

Dose2_VC <- filter(tg, dose == 2 & supp =="VC")
Dose2_OJ <- filter(tg, dose == 2 & supp == "OJ" & len <= 29)

CleanTGData <- rbind(Dose0.5Filter, Dose1_VC, Dose1_OJ, Dose2_VC, Dose2_OJ)

# View(CleanTGData)
ggplot(CleanTGData, aes(x=as.factor(dose), y=len, fill=supp)) + geom_boxplot() + labs(title="Tooth Growth by Dose and Supplement Type (Without Outliers)", x="Dose (mg/day)", y="Tooth Length (microns)", fill="Supplement")
```

This graph shows the same data as the previous graph except in this one, three outliers were removed to provide a more clear analysis.  
<br>
<br>

::: {style="display: flex; gap: 20px;"}
```{r, echo=FALSE}
# This r-block is going to be my sort.
# I will be sorting all of the data points in ToothGrowth in ascending and descending order of the Tooth Length.

AscendingOrder <- arrange(tg, len)
DescendingOrder <- arrange(tg, desc(len))

# Table 1: Shortest
kable(AscendingOrder, row.names = FALSE, caption = "Shortest Teeth First") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE) %>%
  scroll_box(width = "100%", height = "400px")

# Table 2: Longest
kable(DescendingOrder, row.names = FALSE, caption = "Longest Teeth First") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE) %>%
  scroll_box(width = "100%", height = "400px")

```
:::
<div style="clear: both;"></div>
<br>
<br>
```{r}
# This r-block is going to be my aggregation.
# I will be taking the average of all the OJ and all of the VC and comparing the two.

OJ_Filter <- filter(tg, supp == "OJ") 
VC_Filter <- filter(tg, supp =="VC")

OJ_Average <- mean(OJ_Filter$len)
VC_Average <- mean(VC_Filter$len)

Total_Averages <- data.frame(supp = c("OJ", "VC"), average_len = c(OJ_Average, VC_Average))

ggplot(Total_Averages, aes(x = supp, y = average_len, fill = supp)) + geom_col() + labs(title = "Average Tooth Growth by Supplement", x = "Supplement Type", y = "Average Tooth Length (microns)", fill="Supplement")
```
<br>
<br>
  Here we can see that organge juice (OJ) is on average a better choice for tooth growth.  
  While both supplement types are good, if you were in a pinch, I would recommend using orange juice.
<br>
<br>

